---
title: "Housing Price Prediction Data Story"
author: "Jason Wilcox"
date: "July 6, 2018"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r Script, eval=TRUE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(ggplot2)
library(dplyr)
library(corrplot)
library(scales)
library(ggrepel)
library(tidyr)
library(gridExtra)
library(RColorBrewer)
library(forcats) #used for categorical values
library(VIM) #kNN impute


train <- read.csv("C:/Users/jason/Desktop/Capstone/Final Project Housing Prices/traindata.csv")
test <- read.csv("C:/Users/jason/Desktop/Capstone/Final Project Housing Prices/testdata.csv")

test$SalePrice <- NA
combined <- rbind(train, test)
```

## Introduction

Housing prices are always changing and anyone who has bought or sold a home knows that the true value can differ greatly depending on a large number of factors. Knowing which traits and qualities can add or detract from a home's value could help in not only determining the true value, but also aid potential buyers and sellers. Analyzing housing data can grant exceptional insight to the best way a seller could increase the value of the house they have put up for sale. Increasing square footage or the size of the garage may be easy ways for a seller to raise the value before the house does sell. Potential buyers can better understand the actual value of a house before they decide to purchase one and know if they are getting a good deal. Predicting the sales price of different homes is possible with the data set sourced from Kaggle (<https://www.kaggle.com/c/house-prices-advanced-regression-techniques>).

## Data

The dataset comprises of 2919 houses having 80 variables (81 initially). These variables are: 
```{r Data Exploration, eval=TRUE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#                                     Data Exploration
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# _______Sale Price__________
ggplot(combined, aes(x = SalePrice)) +
  geom_histogram(fill = "red", binwidth = 10000) + 
  stat_bin(binwidth = 10000, geom = "text", aes(label = ..count..), vjust = -1.5) +
  scale_x_continuous(limits = c(1, 400000), labels = scales::comma)

summary(combined$SalePrice)
#   Min. 1st Qu.  Median   Mean    3rd Qu.    Max. 
# 34900  129975   163000   180921  214000     755000

# _______SalePrice vs Square Foot______
# Is there a relation to size of house and price?    
ggplot(combined, aes(x = X1stFlrSF + X2ndFlrSF + BsmtFinSF1, y = SalePrice)) +
  geom_point() +
  scale_y_continuous(labels = scales::comma) +
  geom_smooth(method = "lm") +
  scale_x_continuous(name = "Bsmt + 1st + 2nd Floor (sqft)")
# Yes positive correlation between house size and price
# look at the outliers
size <- combined %>% as_tibble() %>% mutate(
  sqft = X1stFlrSF + X2ndFlrSF + BsmtFinSF1
)
# Houses 1299 (11286sqft), 255(6936sqft), 524(6572sqft)
# View(size[c(1299, 524, 1183), ])
# Determine why they are so low in value?


# ______Correlation_______
# Important Variables, using correlation can determine which have big impact
numbers <- which(sapply(combined, is.numeric))
numNames <- names(numbers)

combined_numVar <- combined[, numbers]
cor_numVar <- cor(combined_numVar, use = "pairwise.complete.obs") #find correlation values between all numbered variables
cor_sorted <- as.matrix(sort(cor_numVar[, "SalePrice"], decreasing = TRUE))
#SalePrice      1.00000000
#OverallQual    0.79098160
#GrLivArea      0.70862448
#GarageCars     0.64040920
#GarageArea     0.62343144
#TotalBsmtSF    0.61358055
#X1stFlrSF      0.60585218
#FullBath       0.56066376
#TotRmsAbvGrd   0.53372316
#YearBuilt      0.52289733
#YearRemodAdd   0.50710097

# Top 10, all above 0.5 corr
CorHigh <- names(which(apply(cor_sorted, 1, function(x) abs(x) > 0.5)))
cor_numVar <- cor_numVar[CorHigh, CorHigh]

corrplot.mixed(cor_numVar, tl.col = "black", tl.pos = "lt")
# Shows that there could be issues with colinearity (such as with garagecars and garage area
# both of which have a strong correlation with SalePrice). 

# OverallQual has the highest correlation with SalePrice (0.79). It is a value 1-10 showing the quality of a home
# Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# 1.000   5.000   6.000   6.089   7.000  10.000


# _________Overall Quality vs SalePrice _____________
Quality_Box <- ggplot(combined, aes(x = factor(OverallQual), y = SalePrice)) + # factor() splits graph by numbered value
  geom_boxplot(col = "blue") +
  labs(x = "Overall Quality") +
  scale_y_continuous(labels = comma)
  
Quality_Box + geom_smooth(methond = "lm", se = FALSE, aes(group = 1), color = "red")

# Clearly shows that as Overal Quality goes up, so does Sale Price (not surprising)
# Seems to be a change in rate between 6 and 7. The price/quality point seems to increase there 
# As quality goes up, so does the variance in prices (expected), as features vary and so does associated costs
# No major outliers other than a high cost house with quality 4(mistake?), and possibly a low cost home at quality 10.


#_________Above Ground Living Area vs SalePrice____________
ggplot(combined, aes(x = GrLivArea, y = SalePrice)) +
  geom_point() +
  geom_smooth(method = "lm") + #linear model
  scale_y_continuous(labels = comma)

# Some outliers at 4700 sqft and 5600 sqft to be considered later


#__________Garage Cars vs SalePrice_________
# Size of garage in car capacity

Garage_Box <- ggplot(combined, aes(x = factor(GarageCars), y = SalePrice)) +
  geom_boxplot()
Garage_Box + geom_smooth(method = "lm", se = FALSE, aes(group = 1))
# As garage capacity (in terms of cars) goes up, so would the value of the house with it. 
cat("There are", length(which(combined$GarageCars == 3)), "homes with a 3 car garage")
cat("There are", length(which(combined$GarageCars == 4)), "homes with a 4 car garage")
# Huge difference in the number of homes with a 3 car vs 4 car garage.

#_____________GarageArea vs SalePrice________________
# Size of garage in sqft
ggplot(combined, aes(x = GarageArea, y = SalePrice)) +
  geom_point() +
  geom_smooth(method = "lm")




#---------------------------------------------------
#_____________Factored Categories___________________
factors <- which(sapply(combined, is.factor))


#_____________Price and Others grouped by Neighborhood_________________
# Are some houses more costly based on certain neighborhoods?
citygroup <- combined %>%
  select(Neighborhood, SalePrice, LotArea, LotFrontage, X1stFlrSF, X2ndFlrSF, OverallQual, YearBuilt) %>%
  dplyr::group_by(Neighborhood) %>%
  dplyr::summarize(SalePrice = mean(SalePrice, na.rm = TRUE), LotArea = mean(LotArea, na.rm = TRUE), LotFrontage = mean(LotFrontage, na.rm = TRUE), SF = mean(X1stFlrSF + X2ndFlrSF), OverallQual = mean(OverallQual), YearBuilt = mean(YearBuilt))


# Reorder factor levels of Neighborhood by Saleprice
citygroup$Neighborhood <- factor(citygroup$Neighborhood, levels = citygroup$Neighborhood[order(citygroup$SalePrice)]) 

colorcount <- length(citygroup$Neighborhood) # n count for color pallete
getPalette <- colorRampPalette(brewer.pal(9, "Set1")) # interpolating colors

# Neighborhood by Sale Price
neighborhood1 <- ggplot(citygroup) +
  geom_col(aes(factor(Neighborhood), fill = factor(Neighborhood), SalePrice)) +
  scale_y_continuous(labels = comma) +
  scale_fill_manual(values = getPalette(colorcount), guide = FALSE) +
  geom_hline(yintercept = mean(citygroup$SalePrice), color = "black", linetype = "dashed") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

neighborhood1 # Clearly some of the neighborhoods have much more valuable houses than others
# NoRidge, NridgHt, stoneBr

# Neighborhood by OverallQual
neighborhood2 <- ggplot(citygroup) +
  geom_col(aes(factor(Neighborhood), fill = factor(Neighborhood), OverallQual)) +
  scale_fill_manual(values = getPalette(colorcount), guide = FALSE) +
  scale_y_continuous(limits = c(0, 10)) +
  geom_hline(yintercept = mean(citygroup$OverallQual), color = "black", linetype = "dashed") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

neighborhood2 # Overall similar pattern between OverallQual and SalePrice for each Neighborhood
# Those that rate houses seem to be consistent vs SalesPrice

# Price by OverallQuality
neighborhood3 <- ggplot(citygroup, aes(x = OverallQual, y = SalePrice)) +
  geom_smooth(method = "lm") +
  scale_y_continuous(labels = comma)

neighborhood3
# Price appears to be directly related to OverallQual (shows the people are accurate in their ratings of a house)

# Neighborhood by house age
neighborhood4 <- ggplot(citygroup) +
  geom_col(aes(factor(Neighborhood), fill = factor(Neighborhood), y = (2018 - YearBuilt))) +
  scale_fill_manual(values = getPalette(colorcount), guide = FALSE) +
  scale_y_continuous(name = "Avg. House Age (yrs)") +
  geom_hline(yintercept = mean(2018 - citygroup$YearBuilt), color = "black", linetype = "dashed") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
neighborhood4

# House size by Neighborhood
neighborhood5 <- ggplot(citygroup) +
  geom_col(aes(factor(Neighborhood), fill = factor(Neighborhood), y = (citygroup$SF))) +
  scale_fill_manual(values = getPalette(colorcount), guide = FALSE) +
  scale_y_continuous(labels = comma, name = "Square Footage (sqft)") +
  geom_hline(yintercept = mean(citygroup$SF), color = "black", linetype = "dashed") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

neighborhood5 # Most neighborhoods follow the trend larger homes = higher sales price

# Lot area by Neighborhood
neighborhood6 <- ggplot(citygroup) +
  geom_col(aes(factor(Neighborhood), fill = factor(Neighborhood), y = (citygroup$LotArea))) +
  scale_fill_manual(values = getPalette(colorcount), guide = FALSE) +
  scale_y_continuous(name = "Lot Area (sqft)") +
  geom_hline(yintercept = mean(citygroup$LotArea), color = "black", linetype = "dashed") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


neighborhood6 # Some neighborhoods have significantly larger lot areas (and several significantly less, probably town houses etc)

# Comparison graph
# Neighborhood 1 and 2 compared, Shows some large differences between SalePrice and OverallQual by Neighborhood.
# NoRidge has highest SalePrice but 3rd highest Quality rating, and Blueste has a significantly lower SalePrice vs quality rating
# Can see that places like BrDale, Blueste, NPkVill all have higher quality than their respective neighbors for saleprice. This shows that they are best value
grid.arrange(neighborhood1, neighborhood2, neighborhood5, neighborhood6, neighborhood4, nrow = 5)

grid.arrange(neighborhood1, neighborhood2, nrow =2)
grid.arrange(neighborhood5, neighborhood6, nrow=2)
grid.arrange(neighborhood4, nrow=2)


# ******After adjust the theme on previous graphs, takes up too much space in grid ----> find a solution
```


```{r echo = FALSE, comment=NA}
names(combined)
```

Most are pretty obvious, but some are less clear. Condition 1 and 2 represent the houses proximity to various conditions such as railroads, parks, arterial street, etc. OverallQual is a rating given to the house based on material and finish. BsmtExposure refers to walkout or garden level walls. GrLivArea is the above ground living area in sqft.Further information about the details of each variable can be found on the Kaggle website (<https://www.kaggle.com/c/house-prices-advanced-regression-techniques/data>). The data_description.txt file gives a quick idea of what each variable represents and the possible levels if it is a factor variable. 


## Data Wrangling

Initially the raw test and train data were loaded as a CSV file and bound together to create a full dataset of 2919 observations. I then went systematically through the variables looking for missing values to replace and any other erroneous data points that needed to be addressed. Initially there were 15424 NA's among 37 variables (including the 1459 missing from SalePrice). 

For many categories, it was simply easiest and best suited to just replace any NA's with the most common occuring value. This was done with MSZoning [4], Functional [2], Exterior1st and 2nd [1, 1], Electrical [1], KitchenQual [1], and SaleType [1]. This was done like this:
```{r  Replace most frequent, eval=FALSE, echo= T}
combined$MSZoning[is.na(combined$MSZoning)] <- names(sort(-table(combined$MSZoning)))[1] #replace the 4 missing values with most common factor levels
```

Also very common was simply replacing NA with "None" or some default. This was done in MiscFeature [2814], Alley [2721], GarageYrBlt [159] (Na were converted to match HouseYrBlt since these houses appeared to have no garage or a garage built when the house was built), GarageType [157], GarageCond [159], GarageFinish [159], GarageQual [159], BsmtExposure [82], BsmtCond [82], BsmtQual [81], and MasVnrArea [23]. BsmtFinType 1 and 2 [79, 80], BsmtUnfSF, TotalBsmtSF, BsmtFullBath, and BsmtHallfBath [1, 1, 2, 2] were all replaced with 0's.
```{r Replace 0/None, eval=FALSE, echo= T}
combined$MiscFeature <- factor(ifelse(is.na(combined$MiscFeature), "None", paste(combined$MiscFeature)), levels = c(levels(combined$MiscFeature), "None"))
```

Some values had ordinality naturally or needed to be assigned ordinality. NA's were also present and needed to be converted to "None". PoolQC [2909] and FireplaceQual [1420] already had some ordinality but I wanted to convert them from their system of ratings using letters (Fa = Fair, Gd = Good, etc) to that of numbers so I could treat it as numeric. I assigned a value to each rating and then converted the variable from a factor to numeric.

```{r PoolQC/FireplaceQual, eval = F, Echo = T}
print(fct_count(combined$PoolQC)) # show the number and levels appearing
combined$PoolQC <- factor(ifelse(is.na(combined$PoolQC), "None", paste(combined$PoolQC)), levels = c(levels(combined$PoolQC), "None"))
# Converts the level from NA to None and adds the level None

combined$PoolQC <- factor(combined$PoolQC, levels = c("None", "Po", "Fa", "TA", "Gd", "Ex")) #Reorder levels in ascending order

# Ordinal Values for use later with other variables as well
Quality <- c("None" = 0, "Po" = 1, "Fa" = 2, "TA" = 3, "Gd" = 4, "Ex" = 5)

combined$PoolQC <- (plyr::revalue(combined$PoolQC, Quality)) 
table(combined$PoolQC)

combined$PoolQC <- as.numeric(levels(combined$PoolQC))[combined$PoolQC]
```
Others showed ordinality but didn't use this same system. MasVnrType [24] is an example where I needed to check ordinality. It was clear that stone was better that brick face, which was still better than nothing. I then assigned these levels order and converted to integers. I did something similar with Street and Alley (they had no missing values), where I needed to assign ordinality due to paved roads being significantly better than gravel and dirt roads.

```{r Assigning Ordinality, eval = F, echo = T}
# View(combined[is.na(combined$MasVnrType), c("MasVnrType", "MasVnrArea", "YearBuilt", "ExterCond")])
# House 2611 missing MasVnrType. Impute with most common 
combined$MasVnrType[2611] <- names(sort(-table(combined$MasVnrType)))[2] #most common is none, so chose 2nd most common

# Determine Ordinality for the 23 houses without MasVnrType missing and fill in with None
combined$MasVnrType[is.na(combined$MasVnrType)] <- "None"
exterier <- combined[!is.na(combined$SalePrice),] %>% group_by(MasVnrType) %>% dplyr::summarise(median = median(SalePrice), counts = n()) %>%arrange(median)

print(exterier)

# Clearly ordinal, Stone > BrkFace = None
masonry <- c("None" = 0, "BrkCmn" = 0, "BrkFace" = 1, "Stone" = 2)
combined$MasVnrType<- factor(combined$MasVnrType, levels = c("None", "BrkCmn", "BrkFace", "Stone")) #Reorder levels
combined$MasVnrType <- as.integer(plyr::revalue(combined$MasVnrType, masonry))
table(combined$MasVnrType) #increases the value of everything by 1 why?
```

For LotFrontage [486], GarageCars and GarageArea [1, 1] I used Knn nearest neighbor with a k value of 5 to impute the missing values. This was done after grouping by neighborhood as homes in the same neighborhoods would have similar lot sizes and though it best to fill in the missing values based around that similarity.

```{r Knn Impute, eval = F, echo = T}
impute <- kNN(combined, k = 5)

imputeNeighborhood <- impute %>% select(Neighborhood, LotFrontage) %>%
  dplyr::group_by(Neighborhood) %>% 
  dplyr::summarize(LotFrontage = mean(LotFrontage))

imputeNeighborhood

combined$LotFrontage <- impute$LotFrontage
```

Lastly miscellaneous odds and ends were tied up. For PoolArea [3] I graphed the data and extrapolated to fill in the three missing values. I dropped Utilities since nearly all the values were the same, it provided no useful information. There was a typo I noticed with GarageYrBlt where there was a value of 2207 which I changed back to 2007.

```{r Misc, eval = F}
test <- combined %>% subset(PoolQC > 0)

ggplot(test, aes(x = OverallQual, y = PoolQC)) + 
  geom_point() + 
  geom_smooth(method = "lm") +
  scale_x_continuous(limits = c(0, 10), labels = scales::comma)

combined$PoolQC[2600] <- 2
combined$PoolQC[2421] <- 2
combined$PoolQC[2504] <- 3
```

# Exploratory Data Analysis
```{r Data Wrangling, eval=TRUE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#                                                  Data Wrangling
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

aggr(combined)# Shows all the missing values


NAcol <- which(colSums(is.na(combined)) > 0)
sort(colSums(sapply(combined[NAcol], is.na)), decreasing = TRUE)
# 1459 in SalePrice are the missing values to predict
# PoolQC has the most missing values at 2909
# Ex = Excellent, Gd = Good, TA = Avg, Fa = Fair, NA = no pool


# __________________PoolQC____________________
print(fct_count(combined$PoolQC)) # show the number and levels appearing
combined$PoolQC <- factor(ifelse(is.na(combined$PoolQC), "None", paste(combined$PoolQC)), levels = c(levels(combined$PoolQC), "None"))
# Converts the level from NA to None and adds the level None

combined$PoolQC <- factor(combined$PoolQC, levels = c("None", "Po", "Fa", "TA", "Gd", "Ex")) #Reorder levels

# Ordinal Values for use later with other variables as well
Quality <- c("None" = 0, "Po" = 1, "Fa" = 2, "TA" = 3, "Gd" = 4, "Ex" = 5)

combined$PoolQC <- (plyr::revalue(combined$PoolQC, Quality)) 
table(combined$PoolQC)

combined$PoolQC <- as.numeric(levels(combined$PoolQC))[combined$PoolQC] #says cooerces to NA, but none created?


#__________________PoolArea_____________________

combined[combined$PoolArea > 0, c('PoolArea', 'PoolQC', 'OverallQual')] # See the values for the few pools and if there was a correlation between poolqc, area, and overallqual
ggplot(combined, aes(x = PoolArea, y = PoolQC)) + geom_point() #very little relation
# 2600 and 2421 are similar with OverallQual and will be imputed based on values nearby with similar Overall Quality
# 2504 will be done similarly, with a higher value based on nearby ones

test <- combined %>% subset(PoolQC > 0)

ggplot(test, aes(x = OverallQual, y = PoolQC)) + 
  geom_point() + 
  geom_smooth(method = "lm") +
  scale_x_continuous(limits = c(0, 10), labels = scales::comma)

combined$PoolQC[2600] <- 2
combined$PoolQC[2421] <- 2
combined$PoolQC[2504] <- 3

#______MiscFeature______
combined$MiscFeature <- factor(ifelse(is.na(combined$MiscFeature), "None", paste(combined$MiscFeature)), levels = c(levels(combined$MiscFeature), "None"))
# change NA to None

ggplot(combined, aes(x = MiscFeature, y = SalePrice)) +
  geom_bar(stat = "Summary", fun.y = "median", fill = 'purple') +
  scale_y_continuous(labels = comma) +
  geom_label(stat = "count", aes(label = ..count.., y = ..count..))
# Avg SalePrice by Misc. Feature. Tenis court seems to add the most value (but only 1 house has that feature)
# Houses with a shed have a lower value than those with nothing, which probably means that in place of a garage, the houses have a shed
# No imputation needed for houses with none.

#______Alley______
combined$Alley <- factor(ifelse(is.na(combined$Alley), "None", paste(combined$Alley)), levels = c(levels(combined$Alley), "None"))
# Converts NA's to None

ggplot(combined, aes(x = Alley, y = SalePrice)) +
  geom_bar(stat = "Summary", fill = 'blue', fun.y = "median")
# As expected, houses with a paved alley have higher value than those with gravel ones, and those without only slightly lower.
# No imputation needed

#______Fence______
# 2348 NA for No Fence
combined$Fence <- factor(ifelse(is.na(combined$Fence), "None", paste(combined$Fence)), levels = c(levels(combined$Fence), "None"))

ggplot(combined, aes(x = Fence, y = SalePrice)) +
  geom_bar(stat = "Summary", fill = "dark green", fun.y = "median")

#Doesn't seem to be much difference between no fence and the different values of fences (good privacy, min privacy, good wood, minimum wood/wire, and none)

#______Fireplace Quality______
# 1420 NA for None, which is the same amount as houses with 0 fireplaces, making for easy replacing NA's
combined$FireplaceQu <- factor(ifelse(is.na(combined$FireplaceQu), "None", paste(combined$FireplaceQu)), levels = c(levels(combined$FireplaceQu), "None"))

combined$FireplaceQu <- factor(combined$FireplaceQu, levels = c("None", "Po", "Fa", "TA", "Gd", "Ex")) #Reorder levels
combined$FireplaceQu <- as.integer(plyr::revalue(combined$FireplaceQu, Quality))
table(combined$FireplaceQu)

ggplot(combined, aes(x = FireplaceQu, y = SalePrice)) +
  geom_bar(stat = "summary", fun.y = "median", fill = 'orange')
# SalePrice is on par with quality levels. None is about as equal as a "poor" fireplace

#______Lot Frontage_____
# 486 NA

ggplot(combined[!is.na(combined$LotFrontage),], aes(x = as.factor(Neighborhood), y = LotFrontage)) +
  geom_bar(stat = "summary", fun.y = "median", fill = "blue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

neighborhood7 <- ggplot(citygroup) +
  geom_col(aes(factor(Neighborhood), fill = factor(Neighborhood), y = (citygroup$LotFrontage)))  +
  scale_fill_manual(values = getPalette(colorcount), guide = FALSE) +
  scale_y_continuous(name = "Lot Frontage (sqft)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

neighborhood7
# No matching trend to the SalesPrice based on neighbhorhood, but certainly variance among some. Easiest solution is to use knn to determine based on neighborhoods
citygroup[,c("Neighborhood", "LotFrontage")]

impute <- kNN(combined, k = 5)

imputeNeighborhood <- impute %>% select(Neighborhood, LotFrontage) %>%
  dplyr::group_by(Neighborhood) %>% 
  dplyr::summarize(LotFrontage = mean(LotFrontage))

imputeNeighborhood

combined$LotFrontage <- impute$LotFrontage

#_______GarageYrBlt_________
#159 NA

garage <- combined[,c("GarageType", "GarageYrBlt", "GarageFinish", "GarageCars", "GarageArea", "GarageQual", "GarageCond", "SalePrice")]
# View(combined[is.na(combined$GarageType), c("GarageType", "GarageYrBlt", "GarageFinish", "GarageCars", "GarageArea", "GarageQual", "GarageCond", "SalePrice")])


#GarageYrBlt NA to default to year house was built, similarly with how remodel does with house yr built
for(i in 1:nrow(combined)){
  if (is.na(combined$GarageYrBlt[i])){
    combined$GarageYrBlt[i] <- as.integer(combined$YearBuilt[i])
  }
}

#_______The 2 GarageType with other missing NA' stats___________
# 2127 Has a shed, missing Qual, Condition, and Finish
# 2577 doesn't appear to have a garage, and missing Qual, Condition, and Finish
combined$GarageFinish[2127] <- names(sort(-table(combined$GarageFinish[1])))
combined$GarageQual[2127] <- names(sort(-table(combined$GarageQual[1])))
combined$GarageCond[2127] <- names(sort(-table(combined$GarageCond[1])))

# Check fixed garage info for House 2127
combined[2127, c("GarageYrBlt", "GarageCars", "GarageArea", "GarageType", "GarageCond", "GarageQual", "GarageFinish")]

# Fixing 2577 house values
combined$GarageCars[2577] <- 0
combined$GarageArea[2577] <- 0
combined$GarageType[2577] <- NA #fix the mistakenly labeled Dtached garage so in next step its converted to None

#_______Remaining values without a garage______
#_______GarageType_________
#GarageType NA to None
combined$GarageType <- factor(ifelse(is.na(combined$GarageType), "None", paste(combined$GarageType)), levels = c(levels(combined$GarageType), "None"))

#_______GarageCond/Finish/Qual_______
combined$GarageCond <- factor(ifelse(is.na(combined$GarageCond), "None", paste(combined$GarageCond)), levels = c(levels(combined$GarageCond), "None"))
combined$GarageFinish <- factor(ifelse(is.na(combined$GarageFinish), "None", paste(combined$GarageFinish)), levels = c(levels(combined$GarageFinish), "None"))
combined$GarageQual <- factor(ifelse(is.na(combined$GarageQual), "None", paste (combined$GarageQual)), levels = c(levels(combined$GarageQual), "None"))
# Add "None" to levels and replace NA with None for Cond/Qual/Finish for the houses that don't have a garage.

#______Basement_______
combined[!is.na(combined$BsmtFinType1) & (is.na(combined$BsmtCond)|is.na(combined$BsmtQual)|is.na(combined$BsmtExposure)|is.na(combined$BsmtFinType2)), c("BsmtQual", "BsmtCond", "BsmtExposure", "BsmtFinType1", "BsmtFinType2")]
# Shows missing values where it appears to have basements

# Imputing for the missing values with most common:
combined$BsmtFinType2[333] <- names(sort(-table(combined$BsmtFinType2)))[1]
combined$BsmtExposure[c(949, 1488, 2349)] <- names(sort(-table(combined$BsmtExposure)))[1]
combined$BsmtCond[c(2041, 2186, 2525)] <- names(sort(-table(combined$BsmtCond)))[1]
combined$BsmtQual[c(2218, 2219)] <- names(sort(-table(combined$BsmtQual)))[1]

# Now # of missing values for basement related variables is all the same, representing no basement.
combined$BsmtQual <- factor(ifelse(is.na(combined$BsmtQual), "None", paste(combined$BsmtQual)), levels = c(levels(combined$BsmtQual), "None"))
combined$BsmtCond <- factor(ifelse(is.na(combined$BsmtCond), "None", paste(combined$BsmtCond)), levels = c(levels(combined$BsmtCond), "None"))
combined$BsmtExposure <- factor(ifelse(is.na(combined$BsmtExposure), "None", paste(combined$BsmtExposure)), levels = c(levels(combined$BsmtExposure), "None"))
combined$BsmtFinType1 <- factor(ifelse(is.na(combined$BsmtFinType1), "None", paste(combined$BsmtFinType1)), levels = c(levels(combined$BsmtFinType1), "None"))
combined$BsmtFinType2 <- factor(ifelse(is.na(combined$BsmtFinType2), "None", paste(combined$BsmtFinType2)), levels = c(levels(combined$BsmtFinType2), "None"))

#__________House Masonry_____________

# View(combined[is.na(combined$MasVnrType), c("MasVnrType", "MasVnrArea", "YearBuilt", "ExterCond")])
# House 2611 missing MasVnrType. Impute with most common 
combined$MasVnrType[2611] <- names(sort(-table(combined$MasVnrType)))[2] #most common is none, so chose 2nd most common

# Determine Ordinality for the 23 houses without MasVnrType missing and fill in with None
combined$MasVnrType[is.na(combined$MasVnrType)] <- "None"
exterier <- combined[!is.na(combined$SalePrice),] %>% group_by(MasVnrType) %>% dplyr::summarise(median = median(SalePrice), counts = n()) %>%arrange(median)

# Clearly ordinal, Stone > BrkFace = None
masonry <- c("None" = 0, "BrkCmn" = 0, "BrkFace" = 1, "Stone" = 2)
combined$MasVnrType<- factor(combined$MasVnrType, levels = c("None", "BrkCmn", "BrkFace", "Stone")) #Reorder levels
combined$MasVnrType <- as.integer(plyr::revalue(combined$MasVnrType, masonry))
table(combined$MasVnrType) #increases the value of everything by 1 why?

#______________MasVnrArea____________
# Just needs to be replaced with 0's as the NA houses don't have a vaneer
combined$MasVnrArea[is.na(combined$MasVnrArea)] <- 0

#_____________MSZoning_______________
# 4 value missing

combined$MSZoning[is.na(combined$MSZoning)] <- names(sort(-table(combined$MSZoning)))[1] #replace the 4 missing values with most common factor level
table(combined$MSZoning)

#_____________Utilities______________
# 2 values missing
combined$Utilities[is.na(combined$Utilities)]
table(combined$Utilities) #Since 99% of the values are all AllPub, variable has no use for predicting, so dropping

combined$Utilities <- NULL

#____________Functional____________
# 2 missing values
table(combined$Functional) #Most houses are Typical, no house below 

# imputing with most common: Type
combined$Functional[is.na(combined$Functional)] <- names(sort(-table(combined$Functional)))[1]
 
print(fct_count(combined$Functional))
combined$Functional <- factor(combined$Functional, levels = c("Sal", "Sev", "Maj2", "Maj1", "Mod", "Min2", "Min1", "Typ")) #Reorder levels
Funct <- c("Sal" = 0, "Sev" = 1, "Maj2" = 2, "Maj1" = 3, "Mod" = 4, "Min2" = 5, "Min1" =6, "Typ" = 7)

combined$Functional <- plyr::revalue(combined$Functional, Funct)
table(combined$Functional)

combined$Functional <- as.numeric(levels(combined$Functional))[combined$Functional] #convert to numeric


#________Bsmt Baths/Fin_________
#Several basement variables have 1 or 2 values missing

combined[is.na(combined$BsmtFullBath)|is.na(combined$BsmtHalfBath)|is.na(combined$BsmtFinSF1)|
           is.na(combined$BsmtFinSF2)|is.na(combined$BsmtUnfSF)|is.na(combined$TotalBsmtSF), c("BsmtQual", "BsmtFullBath", 
           "BsmtHalfBath", "BsmtFinSF1", "BsmtFinSF2", "BsmtUnfSF", "TotalBsmtSF")]

# looks like these missing values are for errors, houses without basements. All int type so no need for dealing with factor levels
combined$BsmtFinSF1[2121] <- 0
combined$BsmtFinSF2[2121] <- 0
combined$BsmtUnfSF[2121] <- 0
combined$TotalBsmtSF[2121] <- 0
combined$BsmtFullBath[2121] <- 0
combined$BsmtHalfBath[2121] <- 0

combined$BsmtFinSF1[2189] <- 0
combined$BsmtFinSF2[2189] <- 0
combined$BsmtFullBath[2189] <- 0
combined$BsmtHalfBath[2189] <- 0


#_______________Exterior___________________
combined[is.na(combined$Exterior1st)|is.na(combined$Exterior2nd),]
#Missing exterior 1st and 2nd, impute with most common value
combined$Exterior1st[is.na(combined$Exterior1st)] <- names(sort(-table(combined$Exterior1st)))[1]
combined$Exterior2nd[is.na(combined$Exterior2nd)] <- names(sort(-table(combined$Exterior2nd)))[1]

#___________________Electrical________________
combined[is.na(combined$Electrical),] #1380 missing value
#impute with most common value
combined$Electrical[is.na(combined$Electrical)] <- names(sort(-table(combined$Electrical)))[1]

#__________________KitchenQual________________
combined[is.na(combined$KitchenQual),] #1556 missing value, impute with most common (TA)
combined$KitchenQual[is.na(combined$KitchenQual)] <- names(sort(-table(combined$KitchenQual)))[1]

#_________________GarageCars/Area_________________
combined[is.na(combined$GarageCars), c("GarageType", "GarageYrBlt", "GarageCars", "GarageArea")] #2577 missing both garage cars and area. Appears to have detatched
combined[is.na(combined$GarageArea), ] #2577

impute <- kNN(combined, k = 5) #impute values for garagearea and number of cars

combined[2577, c("GarageCars", "GarageArea")] <- impute[2577, c("GarageCars", "GarageArea")] #replace with imputed values

#_________________SaleType_______________
combined[is.na(combined$SaleType), ] #2490 missing, replace with most common WD

combined[2490, "SaleType"] <- "WD"

#_____________TYPOS______________
#combined$GarageYrBlt has a typo. House #2593 has a yr built as 2207, clearly to be 2007
combined$GarageYrBlt[2593] <- 2007


#___________Ordinal Values_________
combined$Street <- as.integer(plyr::revalue(combined$Street, c("Grvl" = 0, "Pave" = 1)))

combined$PavedDrive <- as.integer(plyr::revalue(combined$PavedDrive, c("N" = 0, "P" = 1, "Y" = 2)))



#___________INT -> Factors__________
#MSSubclass

combined$MSSubClass <- as.factor(combined$MSSubClass)
```

## SalePrice

Having thousands of data points on a diverse group of houses, there are many interesting data points to discuss. There were some very obvious ways to initially look at the data but I also discovered more abstract ways to see it too. Some important findings to my data analysis included simple comparison of sale price with basic things such as house size, age, etc but also some interested things were seen when comparing houses by neighborhoods to see trends across the city.

Initially, the very first thing I wanted to see was sale price shown by count. This gives a quick idea of how the data is represented. As can be seen below, the sale price is roughly normal, skewed to the right slightly which makes sense, considering there isn't an upper limit on housing prices and expensive houses can be built anywhere. There appears to be a noticeable dip in value near where the mean and median area. One goal is to understand why this occurs.Also of note, there are 1459 NA's, which are the missing values to be predicted for SalePrice.

```{r SalePrice Histogram, echo=FALSE, message=FALSE, warning=FALSE}
# _______Sale Price__________
ggplot(combined, aes(x = SalePrice)) +
  geom_histogram(fill = "red", binwidth = 10000) + 
  stat_bin(binwidth = 10000, geom = "text", aes(label = ..count..), vjust = -1.5) +
  scale_x_continuous(limits = c(1, 400000), labels = scales::comma)

summary(combined$SalePrice)
#   Min. 1st Qu.  Median   Mean    3rd Qu.    Max. 
# 34900  129975   163000   180921  214000     755000
```

## House Size

The first predictable analysis I did was sale price vs house size. I used the sum of the basement, 1st, and 2nd floor square footage to create a scatter plot of the various house sizes and what they sold for. As expected, there was a strong trend for the larger the house, the more it sold for. There appear to be a few exceptions (houses 1299 and 524) which appear to have sold for significantly less than expected for the size. I plan on looking into this more to determine the reasons that these houses could be valued so low, perhaps they are old, or the quality is very low. Those were my initial thoughts and a quick look at the data shows they were both recently built (2008 and 2007) and have been given an "OverallQual" value of 10, the highest (and also the most influential variable for SalePrice based on correlation). So it will be interesting to determine why those two houses are much lower in price.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# _______SalePrice vs Square Foot______
# Is there a relation to size of house and price?    
ggplot(combined, aes(x = X1stFlrSF + X2ndFlrSF + BsmtFinSF1, y = SalePrice)) +
  geom_point() +
  scale_y_continuous(labels = scales::comma) +
  geom_smooth(method = "lm") +
  scale_x_continuous(name = "Bsmt + 1st + 2nd Floor (sqft)")
# Yes positive correlation between house size and price
# look at the outliers
size <- combined %>% as_tibble() %>% mutate(
  sqft = X1stFlrSF + X2ndFlrSF + BsmtFinSF1
)
# Houses 1299 (11286sqft), 255(6936sqft), 524(6572sqft)
# View(size[c(1299, 524, 1183), ])
# Determine why they are so low in value?
```

```{r Correlation, echo=FALSE, message=FALSE, warning=FALSE}
corrplot.mixed(cor_numVar, tl.col = "black", tl.pos = "lt")
# Shows that there could be issues with colinearity (such as with garagecars and garage area
# both of which have a strong correlation with SalePrice). 
```

## Overall Quality
Using a correlation plot, I was able to determine that OverallQual had the largest effect on SalePrice, having a correlation value of 0.79. I wanted to better understand just how much impact OverallQual had so I decided to make a graph. I was surprised to actually see that OverallQual which "Rates the overall materiel and finish of the houses", which would also be subjective bias from whoever was giving the ratings, was so strongly related. As quality rating goes up, so does the value of the house (which makes sense). There also appears to be a large change in rate between 6 and 7, suggesting that materials with quality of 7 or higher have greater impact on increasing a home's value per unit.

```{r echo=FALSE, message=FALSE, warning=FALSE}
Quality_Box <- ggplot(combined, aes(x = factor(OverallQual), y = SalePrice)) + # factor() splits graph by numbered value
  geom_boxplot(col = "blue") +
  labs(x = "Overall Quality") +
  scale_y_continuous(labels = comma)
  
Quality_Box + geom_smooth(methond = "lm", se = FALSE, aes(group = 1), color = "red")

# Clearly shows that as Overal Quality goes up, so does Sale Price (not surprising)
# Seems to be a change in rate between 6 and 7. The price/quality point seems to increase there 
# As quality goes up, so does the variance in prices (expected), as features vary and so does associated costs
# No major outliers other than a high cost house with quality 4(mistake?), and possibly a low cost home at quality 10.
```

## Neighborhood

Lastly, to look at things a little differently I decided to group everything by neighborhood to see what trends there were by different areas of the city, and wasn't surprised when I noticed some significant differences. Most people know that cities have both good and bad neighborhoods which can be determined by a number of things, including the type and quality of houses, distance to and quality of schools, hospitals, churches, police, and crime rates. First, I ranked the neighborhoods in ascending order by average SalePrice for the homes for each and marked the mean with a dashed line. I then used the same ascending order to show each neighborhood's OverallQual, SquareFootage, LotArea, and average age of house. Stitching the plots together then allows for simple and easy comparison. Looking at OverallQual, there are some neighborhoods that appear to have houses which have higher quality relative to their SalePrice, such as in Blueste. This means that some homes are selling for less than their true value and would be a great purchase. With the house size graph, some neighborhoods have houses larger than those of similar value, such as in SWISU neighborhood and could provide a large family a lower priced home than potentially elsewhere in the city. Lot Area and house age provided some insight to the other graphs by having less apparent trends due to more outliers but still providing useful information. Neighborhoods that had more expensive homes tend to also have larger plots of land and be newer, but not always.

```{r echo=FALSE, message=FALSE, warning=FALSE}

grid.arrange(neighborhood1, neighborhood2, nrow =2)
grid.arrange(neighborhood5, neighborhood6, nrow=2)
grid.arrange(neighborhood4, nrow=2)

```

# Conclusion

There is quite a lot of data still to be looked at and understand what is going on. I have some questions already to look deeper for the answers and know many more will show up as I dig deeper. There will be more need to analyze the factor variables as well for the influences they may have on the value of homes, which can't be as easily quantified as the initial data was since it was numeric.


 